---
title: "Data_analysis"
author: "Magda Cardozo"
date: "19/11/2019"
output: github_document
---

## Load required libraries

```{r fun, message=TRUE, warning=TRUE}
library(rstatix);packageVersion("rstatix")
library(dplyr);packageVersion("dplyr")
library(tidyr);packageVersion("tidyr")
library(ggsignif);packageVersion("ggsignif")
library(rcompanion);packageVersion("rcompanion")
library(kableExtra);packageVersion("kableExtra")
library(cowplot);packageVersion("cowplot")
library(broom);packageVersion("broom")
library(tidyverse);packageVersion("tidyverse")
library(tidyverse);packageVersion("ggpubr")
```

## Import data

```{r import and preprocess raw counts}
source("import_cell_counts.R")
```

### Bacterioplankton communities strongly change in cell abundance and composition with depth

## Significance of DAPI-stained cells  with depth

```{r sigDAPIDpth}
#test for normal distribution
set.seed(1)
shapiro_DAPI <-shapiro.test(counts_all$DAPI.conc.mn)
shapiro_DAPI
#not normally distributed, therefore applying Kruskal Wallis test
set.seed(2)
kruskal.test(DAPI.conc.mn ~ Depth, data = counts_all)
#posthoc wilcoxon test to compared the different depths, here we observed a significant decrease
set.seed(3)
DAPI_Wilcox <- counts_all  %>%
  wilcox_test(DAPI.conc.mn ~ Depth, p.adjust.method = "bonferroni") %>%
  add_significance()
DAPI_Wilcox
``` 

## Significance of Bacteria/Archaea with depth

#Bacteria

Obtian mean abundaces by region
```{r sigEUBDpth}
counts_all%>% 
  filter(Domain %in% c("EUB")) -> EUB_sum

counts_all %>% filter(Domain %in% c("EUB"))%>%
  group_by(Region, Depth) %>% 
  summarise(mean.EUB = mean(FISH.conc.mn)/1e5,
            sd.EUB =sd(FISH.conc.mn)/1e5)-> Mean_EUB_by_region
```

Test significance with depth
```{r sigEUB}
#test for normal distribution
set.seed(4)
shapiro_EUB <-shapiro.test(EUB_sum$FISH.conc.mn)
shapiro_EUB
#not normally distributed, therefore applying Kruskal Wallis test
set.seed(5)
kruskal.test(FISH.conc.mn ~ Depth, data = EUB_sum)
#posthoc wilcoxon test to compared the different depths, here we observe a signifcant decrease in all depths
set.seed(6)
EUB_Wilcox <- EUB_sum  %>%
  wilcox_test(FISH.conc.mn ~ Depth, p.adjust.method = "bonferroni") %>%
  add_significance()
EUB_Wilcox
```

#Archaea

Obtain mean Archaea abundaces by region
```{r sigARCHDpth}
counts_all%>% 
  filter(Domain %in% c("ARCH")) -> ARCH_sum

counts_all %>% filter(Domain %in% c("ARCH"))%>%
  group_by(Region, Depth) %>% 
  summarise(mean.ARCH = mean(FISH.conc.mn)/1e4,
            sd.ARCH =sd(FISH.conc.mn)/1e4)-> Mean_ARCH_by_region
```

Test significance with depth
```{r sigARCH}
#test for normal distribution
set.seed(7)
shapiro_ARCH <-shapiro.test(ARCH_sum$FISH.conc.mn)
shapiro_ARCH
#not normally distributed, therefore applying Kruskal Wallis test
set.seed(8)
kruskal.test(FISH.conc.mn ~ Depth, data = ARCH_sum)
#posthoc wilcoxon test to compared the different depths, cell abundaces decrease significantly from epi to meso and bathypelagic waters only
set.seed(9)
ARCH_Wilcox <- ARCH_sum  %>%
  wilcox_test(FISH.conc.mn ~ Depth, p.adjust.method = "bonferroni") %>%
  add_significance()
ARCH_Wilcox
```

```{r sumDAPIDpth}
counts_all%>%
  group_by(Region, StationName, Depth) %>%
  summarise(mean.DAPI = mean(DAPI.conc.mn)/1e5,
            sd.DAPI =sd(DAPI.conc.mn)/1e5)-> Total_DAPI_by_Station

counts_all %>%
  group_by(Region, Depth) %>%
  summarise(mean.DAPI = mean(DAPI.conc.mn)/1e5,
            sd.DAPI =sd(DAPI.conc.mn)/1e5)-> Mean_DAPI_by_region
```

## Bacterioplankton cell abundances in the different regions of the Fram Strait (Figure 2.)

Plot
```{r ptVER}
depth_names <- c(`SRF` = "surface (15-30m)",
                 `EPI` = "epipelagic (100m)",
                 `MESO` = "mesopelagic (500-1000m)",
                 `BATHY` = "bathypelagic (>1200m)")

DAPI_vertical_boxplot <- ggplot(counts_all, aes(x= Region, y = DAPI.conc.mn))+
  geom_boxplot(aes(fill = Region))+
  facet_grid(Depth~.,labeller = as_labeller(depth_names))+
  geom_jitter(aes(x= Region, y = DAPI.conc.mn),width = 0.2, alpha = 0.3)+
#  geom_signif(comparisons = list(c("EGC", "WSC"),c("EGC","N"),c("N","WSC")), 
#              map_signif_level=TRUE, test = "wilcox.test")+
  scale_y_log10()+
  ylab(bquote('Cell abundance ('*cells ~mL^-1*')'))+
  scale_fill_manual(values=c("WSC" = "red", "EGC" = "blue", "N" = "gray")) +
  theme_bw()+
  xlab("")+
 # coord_flip()+
  scale_x_discrete(position = "top")+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "none",strip.background =element_rect(fill="white"))

counts_all%>% 
  filter(Domain %in% c("EUB","ARCH"))%>%
  group_by(Region, StationName, Depth, Domain) -> EUB_ARCH_absolute

p <- list()
for (i in c("EUB","ARCH")){
  p[[i]] <-  ggplot(counts_all[counts_all$Domain== i,], aes(x= Region, y = FISH.conc.mn))+
    geom_boxplot(aes(fill = Region))+
    facet_grid(Depth~.,labeller = as_labeller(depth_names))+
    geom_jitter(aes(x= Region, y = FISH.conc.mn),width = 0.2, alpha = 0.3)+
#    geom_signif(comparisons = list(c("EGC", "WSC"),c("EGC","N"),c("N","WSC")), 
#                map_signif_level=TRUE, test = "wilcox.test")+ # significance by region
    scale_y_log10()+
    ylab("")+
    xlab("")+
    #ylab(bquote('Cell abundance ('*cells ~mL^-1*')'))+
    scale_fill_manual(values=c("WSC" = "red", "EGC" = "blue", "N" = "gray")) +
    theme_bw()+
#    coord_flip()+
    scale_x_discrete(position = "top")+ 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "none",strip.background =element_rect(fill="white"))
}

x <- plot_grid(DAPI_vertical_boxplot, p[["EUB"]],p[["ARCH"]], ncol =3, labels = "auto")
x
#tiff("test.tiff", units="cm", width=19, height=17, res=800)
#plot_grid(DAPI_vertical_boxplot, p[["EUB"]],p[["ARCH"]], ncol =3, labels = "auto")
#dev.off()
```

Barplot surface counts
```{r barplot}
taxa_sur_names <- c(`SAR11` = "SAR11 clade",
                 `ROS` = "Rhodobacteraceae",
                 `GAM` = "Gammaproteobacteria",
                 `ALT` = "Alteromonadaceae/Colwelliaceae/Pseudoalteromonadaceae",
                 `BACT` = "Bacteroidetes",
                 `POL` = "Polaribacter",
                 `VER` = "Verrucomicrobiales",
                 `OPI` = "Opitutales")

taxa.sur <- c("SAR11","ROS","GAM","ALT","BACT","POL","VER","OPI")
counts_taxa.sur <- counts_all[counts_all$Domain %in% taxa.sur, ]
counts_Sur <- subset(counts_taxa.sur, Depth == "SRF")
counts_Sur$Domain <- factor(counts_Sur$Domain, levels = c("SAR11","ROS","GAM","ALT","BACT","POL","VER","OPI"))

barplot <- ggplot(counts_Sur, aes(x= StationName, y = FISH.conc.mn, fill = Region))+
  geom_bar(stat = "identity", colour="black")+
  scale_fill_manual(values=c("WSC" = "red", "EGC" = "blue", "N" = "gray")) +
  ylab(bquote('Cell abundance ('*cells ~mL^-1*')'))+
  xlab("")+
  facet_wrap(~ Domain,scales = "free_x", ncol=2,labeller = as_labeller(taxa_sur_names))+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "none",strip.background =element_rect(fill="white"))
barplot

#tiff("Figure-3-surface_counts.tiff", units="cm", width=20, height=23, res=800)
#barplot
#dev.off() 
```

## Coverage of Bacteria and Archaea probes in comparison to DAPI

Obtain proportions for Bacteria and Archaea
```{r covEA}
counts_all%>% 
  filter(Domain %in% c("EUB","ARCH"))%>%
  group_by(Region, StationName, Depth, Domain) %>% 
  summarise(proportion = FISH.conc.mn/DAPI.conc.mn)%>% 
  spread(Domain, proportion)%>%
  mutate(Total.prop = ARCH+EUB) %>%
  ungroup() %>%
  as.data.frame()-> EUB_ARCH_prop
```

Bacteria coverage test
```{r covEUB}
#test for normal distribution
set.seed(10)
shapiro_EUB <-shapiro.test(EUB_ARCH_prop$EUB)# normally distributed p > 0.01 
shapiro_EUB
#ANOVA for depth
set.seed(11)
aov.EUB_Depth <- aov(EUB ~ Depth, data = EUB_ARCH_prop)
summary(aov.EUB_Depth)
TukeyHSD(aov.EUB_Depth) # signifcant decrease from surface to deeper water layers
#ANOVA for region
set.seed(12)
aov.EUB_Region <- aov(EUB ~ Region, data = EUB_ARCH_prop)
summary(aov.EUB_Region) #not significant differences between regions
```

Archaea coverage test
```{r covARCH}
#test for normal distribution
set.seed(13)
shapiro_ARCH <-shapiro.test(EUB_ARCH_prop$ARCH) # normally distributed p > 0.01 
shapiro_ARCH
#ANOVA for depth
set.seed(14)
aov.ARCH_Depth <- aov(ARCH ~ Depth, data = EUB_ARCH_prop)
summary(aov.ARCH_Depth)
TukeyHSD(aov.ARCH_Depth) #significant increase from surface to deeper layers
#ANOVA for region
set.seed(15)
aov.ARCH_Region <- aov(ARCH ~ Region, data = EUB_ARCH_prop)
summary(aov.ARCH_Region) #not significant differences
```

Total coverage
```{r covT}
EUB_ARCH_prop%>%
  group_by(Region, StationName, Depth) %>%
  summarise(total.coverage = sum(Total.prop),
            sd.coverage = sd(Total.prop)) -> EUB_ARCH_coverage
```

By regions
```{r covR}
EUB_ARCH_coverage %>%
  group_by(Region, Depth) %>%
  summarise(mean.coverage = mean(total.coverage),
            sd.coverage = se(total.coverage)) -> EUB_ARCH_coverage_by_region
```

Proportions by depth
```{r covD}
EUB_ARCH_coverage %>%
  group_by(Depth) %>%
  summarise(mean.coverage = mean(total.coverage),
            sd.coverage = se(total.coverage)) -> EUB_ARCH_coverage_by_depth
```            

### Surface water bacterioplankton communities are affected by distinct phytoplankton bloom conditions across the Fram Strait 

## Significance of surface DAPI-stained cells between regions

```{r sigDAPIReg}
#select surface counts
counts_all%>% 
  filter(Depth %in% c("SRF"))  -> DAPI_sum_srf 
##test for normal distribution
set.seed(16)
shapiro_DAPI_srf <-shapiro.test(DAPI_sum_srf$DAPI.conc.mn)# not normally distributed p < 0.01 
shapiro_DAPI_srf
#kruskal wallis test
set.seed(17)
kruskal.test(DAPI.conc.mn ~ Region, data = DAPI_sum_srf)
kruskal.test
set.seed(18)
Region_DAPI_srf_Wilcox <- DAPI_sum_srf   %>%
  rstatix::wilcox_test(DAPI.conc.mn ~ Region, p.adjust.method = "BH") %>%
  add_significance()
Region_DAPI_srf_Wilcox #significant differences between regions
```

## Significance of groups between the regions 

Here we test each group significant differences between regions
```{r sigGrReg}
set.seed(19)
Region_FISH_srf_Wilcox <- DAPI_sum_srf   %>%
  group_by(Domain)%>%
  rstatix::wilcox_test(FISH.conc.mn ~ Region, p.adjust.method = "BH") %>%
  add_significance()
Region_FISH_srf_Wilcox 
Region_FISH_srf_Wilcox_sig <- filter(Region_FISH_srf_Wilcox, p.adj < 0.01)
Region_FISH_srf_Wilcox_sig #no group was significant
```

## Proportional abundance of the different groups at the surface

Obtain proportions of surface taxonomic groups
```{r FISHSRF}
counts_all%>% 
  filter(Depth %in% c("SRF"))%>%
  group_by(Region, StationName, Domain)%>%
  mutate(proportion = (FISH.conc.mn / DAPI.conc.mn)*100) %>%
  select(Region, StationName, Domain, FISH.conc.mn, proportion)  -> surface_FISH_proportion
```

Test significance
```{r sigFISH}
set.seed(20)
Region_FISH_prop_srf_Wilcox <- surface_FISH_proportion   %>%
  group_by(Domain)%>%
  rstatix::wilcox_test(proportion ~ Region, p.adjust.method = "BH") %>%
  add_significance()
Region_FISH_prop_srf_Wilcox
Region_FISH_prop_srf_Wilcox_sig <- filter(Region_FISH_prop_srf_Wilcox, p.adj < 0.01)
Region_FISH_prop_srf_Wilcox_sig #no significant differences
```

## Overview of phylogenetic group abundances by regions 

Calculate abundance in surface by region 
```{r FISH}
DAPI_sum_srf %>%
  group_by(Region, Domain) %>%
  summarise(mean.abund = mean(FISH.conc.mn),
            sd.abund = sd(FISH.conc.mn),
            mean.DAPI = mean(DAPI.conc.mn),
            sd.DAPI = sd(DAPI.conc.mn),
            n= length(FISH.conc.mn))-> FISH_abundance_surface_by_region
```

Make wide table
```{r FISHw}
FISH_abundance_surface_by_region %>%
  select(Region,Domain,mean.abund)%>%
  spread(Region,mean.abund) -> FISH_abundance_surface_by_region_wide
```

Calculate abundances by region
```{r FISHreg}
counts_all%>%
  group_by(Region, Domain, Depth) %>%
  summarise(mean.abund = mean(FISH.conc.mn)/1e5,
            se.abund = se(FISH.conc.mn)/1e5,
            n= sum(n))-> FISH_abundances_by_region

FISH_abundances_by_region$Depth <- factor(FISH_abundances_by_region$Depth, levels = c("SRF","EPI","MESO","BATHY"))
 FISH_abundances_by_region %>%
   select(Region,Domain,Depth,mean.abund)%>%
   spread(Domain,mean.abund) -> FISH_abundance_by_region_wide

#write.csv(FISH_and_dapi_abundance_by_region_w, file="~/CARD-FISH/CARD-FISH_water_project/cell_densities.csv")
```

Summarize by regions
```{r FSRFreg}
surface_FISH_proportion%>%
  group_by(Region, Domain) %>%
  summarise (mean.abund = mean(FISH.conc.mn),
             se.abund = sd(FISH.conc.mn),
             mean.prop = mean(proportion),
             se.prop = sd(proportion)) -> surface_FISH_proportion_by_regions
```

Depth profiles
```{r plot2}
taxa.de <- c("CREN","SAR406","DELTA","SAR324","CFX","SAR202")
counts_taxa.de <- FISH_abundances_by_region[FISH_abundances_by_region$Domain %in% taxa.de, ]
counts_taxa.de$Depth<- factor(counts_taxa.de$Depth, 
                        levels = rev(c("SRF", "EPI", "MESO", "BATHY")))
counts_taxa.de$Domain<- factor(counts_taxa.de$Domain, 
                        levels =c("CREN","SAR406","DELTA","SAR324","CFX","SAR202"))

domain_names <- c(`CREN`="Thaumarchaeota",
                  `SAR406`="SAR406 clade",
                  `DELTA`="Deltaproteobacteria",
                  `SAR324`="SAR324 clade",
                  `CFX`="Chloroflexi",
                  `SAR202`="SAR202 clade")


Plot_depth_profiles.p <- ggplot(data = counts_taxa.de, aes(y = mean.abund, x = Depth))+
  geom_line(data = counts_taxa.de, linetype=1, aes(color =Region, group=Region))+
  facet_grid(.~Domain,labeller=labeller(Domain = domain_names))+
  #geom_errorbarh(data = FISH_abundances_by_region, aes(xmin=mean.abund-se.abund, xmax=mean.abund+se.abund)) +
  scale_x_discrete(labels=c("SRF"="surface (15-30m)", "EPI"="epipelagic (100m)", "MESO"="mesopelagic (500-1000m)","BATHY"="bathypelagic (>1200m)"))+
  xlab("")+
  scale_y_log10()+
  ylab(bquote('Cell abundance ('*10^5*''*cells ~mL^-1*')'))+
    coord_flip()+
  geom_errorbar(data = counts_taxa.de, aes(ymin=mean.abund-se.abund, ymax=mean.abund+se.abund), width=0.09)+
  geom_point(data = counts_taxa.de, aes(y = mean.abund, x = Depth),shape=21, fill="white")+
  scale_color_manual(values=c("WSC" = "red", "EGC" = "blue", "N" = "darkgray")) +
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = "none",strip.background =element_rect(fill="white"))
Plot_depth_profiles.p

# tiff("Figure-4-Depth_profiles.tiff", units="cm", width=33, height=9, res=800)
# Plot_depth_profiles.p
# dev.off() 
```

```{r summarize_master_recyclers}
surface_FISH_MR<- surface_FISH_proportion%>%
  mutate(Master_Rec = if_else(Domain %in% c("BACT","GAM","VER"), "yes","no"))%>%
  group_by(Region, StationName, Master_Rec) %>%
  summarise (sum.abund = sum(FISH.conc.mn),
             sum.prop = sum(proportion))%>%
  filter(Master_Rec =="yes")
```

Calculate proportional abundance of the different groups
```{r FSRFgr}
counts_all%>%
  group_by(Region, StationName, Domain)%>%
  mutate(proportion = (FISH.conc.mn / DAPI.conc.mn)*100) %>%
  group_by(Region, Domain, Depth) %>%
  summarise(mean.prop = mean(proportion),
            se.prop = se(proportion))%>%
  select(Region, Domain, mean.prop, Depth) %>%
  spread(Domain,mean.prop) ->  FISH_proportion

#write.csv(FISH_proportion, file="~/CARD-FISH/CARD-FISH_water_project/cell_prortions.csv")
```  

## Correlation between environmental parameters and cell abundances

Import and normalize environmental parameters
```{r envpar-imp}
#define scale function
scale_par <- function(x) scale(x, center = FALSE, scale = TRUE)[,1]
#import metadata
metadata.raw <- read.csv("~/CARD-FISH/PS99_CARD-FISH_Analysis_final_folder(fill it in the end)/Data/Consumed_nutrients_PS99_2.csv", sep = ",", dec = ".", header = TRUE)
#select variables
env.par <- c("Temperature","Salinity","Chla_conc", "d.NO3.", "d.PO4.", "d.SiO3.", "NH4")

#scale all the parameters and extract surface samples
metadata.scaled.SRF <- metadata.raw %>% 
  mutate_at(env.par, scale_par)%>%
  as.data.frame()

#import protist counts and add to env. table
protist.counts <- read.csv("../Data/PS99_protist_counts.csv",
                       sep = ",", dec = ".", header = TRUE)

metadata.scaled.SRF_p <- inner_join(
  metadata.scaled.SRF[,c("StationName", "Depth.m.", "Temperature","Salinity","Chla_conc","d.NO3.","d.PO4.","d.SiO3.")], 
  protist.counts,
  by = "StationName")
metadata.scaled.SRF_p$Depth.m.<- NULL
metadata.scaled.SRF_p$Depth<- NULL
```

Correlation test between the env. parameters and plot (no protist included)
```{r envpar-selfcorr}
set.seed(21)
envpar_corr <- metadata.scaled.SRF %>% 
  select("Temperature","Salinity","Chla_conc", "d.NO3.", "d.PO4.", "d.SiO3.", "NH4")%>% 
   cor_mat(method = "pearson")
envpar_corr
#check the p-values
envpar_corr.pvalues <- envpar_corr %>% cor_get_pval()

#plot
if(require("ggpubr")){
  my.palette <- get_palette("PuOr", 200)
  envpar_corr %>%
  cor_reorder() %>%
  pull_lower_triangle() %>%
  cor_plot(label = TRUE, palette = my.palette, method = "number")
}

#write output
#write.csv(envpar_corr, "./corr_table_env_env.csv")
#write.csv(envpar_corr.pvalues, "./p_table_env_env.csv")
```

Correlation test between the env. parameters and plot (with protist groups)
```{r envpar-selfcorr_p}
set.seed(22)
envpar_corr.p <- as.data.frame(metadata.scaled.SRF_p) %>% 
  select(-c(StationName))%>% 
   cor_mat(method = "pearson")
envpar_corr.p
#check the p-values
envpar_corr.pvalues.p <- envpar_corr.p %>% cor_get_pval()

#get correlation cooficients
corr_rho.p<- envpar_corr.p %>% select(rowname, "Diatoms", "Phaeocystis.spp","Coccolithophores","Nanoflagellates","Prorocentrum.spp","Myrionecta.rubra","Dinoflagellates","Ciliates","Tintinnids")
#check p-values
corr_p.p <- envpar_corr.p %>% cor_get_pval()%>% select(rowname, "Diatoms", "Phaeocystis.spp","Coccolithophores","Nanoflagellates","Prorocentrum.spp","Myrionecta.rubra","Dinoflagellates","Ciliates","Tintinnids")

#plot
envpar_corr.p %>%
  cor_reorder() %>%
  pull_lower_triangle() %>%
  cor_plot(label = TRUE)

#write output
#write.csv(corr_rho.p, "./corr_table_env_protist.csv")
#write.csv(corr_p.p, "./p_table_env_protist.csv")
```

Correlation between environmental parameters and cell abundances of different groups and plot
```{r envpar-counts-corr}
#merge counts and env.par data
data_4_corr.sub <- counts_all%>% 
  filter(Depth %in% c("SRF"))%>%
  group_by(Region, StationName, Domain)%>%
  select(Region, StationName, Domain, FISH.conc.mn)  %>% 
  spread(Domain, FISH.conc.mn)  %>%
  #drop samples with no env. data
  filter(StationName %in% metadata.scaled.SRF_p$StationName) %>% 
  #merge counts and environmental data
  left_join(., metadata.scaled.SRF_p, by = "StationName")

#Correlation test
set.seed(23)
counts_env_par_corr <- as.data.frame(data_4_corr.sub) %>% 
  select(-c(Region,StationName))%>% 
  cor_mat(method = "pearson")
#get correlation cooficients
corr_rho<- counts_env_par_corr %>% select(rowname,"Temperature","Salinity","Chla_conc","d.NO3.","d.PO4.","d.SiO3.", "Diatoms", "Phaeocystis.spp","Coccolithophores","Nanoflagellates","Prorocentrum.spp","Myrionecta.rubra","Dinoflagellates","Ciliates","Tintinnids")
#check p-values
corr_p <- counts_env_par_corr %>% cor_get_pval()%>% select(rowname,"Temperature","Salinity","Chla_conc","d.NO3.","d.PO4.","d.SiO3.", "Diatoms", "Phaeocystis.spp","Coccolithophores","Nanoflagellates","Prorocentrum.spp","Myrionecta.rubra","Dinoflagellates","Ciliates","Tintinnids")

#plot
counts_env_par_corr %>%
  cor_reorder() %>%
  pull_lower_triangle() %>%
  cor_plot(significant.level = 0.1,label = TRUE)

#write output
#write.csv(corr_rho, "./corr_table_counts_env_protist.csv")
#write.csv(corr_p, "./p_table_counts_env_protist.csv")
```

Correlation between environmental parameters and cell abundances of different groups (no protists)
```{r envpar-counts-corr_nop}
#merge counts and env.par data
data_4_corr <- counts_all%>% 
  filter(Depth %in% c("SRF"))%>%
  group_by(Region, StationName, Domain)%>%
  select(Region, StationName, Domain, FISH.conc.mn)  %>% 
  spread(Domain, FISH.conc.mn)  %>%
  #drop samples with no env. data
  filter(StationName %in% metadata.scaled.SRF$StationName) %>% 
  #merge counts and environmental data
  left_join(., metadata.scaled.SRF[,c("StationName",env.par)] , by = "StationName")

#Correlation test
set.seed(24)
counts_env_par_corr_all <- as.data.frame(data_4_corr) %>% 
  select(-c(Region,StationName))%>% 
  cor_mat(method = "pearson")
#get correlation cooficients
corr_rho.a<- counts_env_par_corr_all %>% select(rowname,env.par)
#check p-values
corr_p.a <- counts_env_par_corr_all %>% cor_get_pval()%>% select(rowname,env.par)

#plot
counts_env_par_corr_all %>%
  cor_reorder() %>%
  pull_lower_triangle() %>%
  cor_plot(significant.level = 0.1,label = TRUE)
```

Correlation between and cell abundances
```{r counts-counts-corr}
#merge counts and env.par data
data_4_corr.c <- counts_all%>% 
  filter(Depth %in% c("SRF"))%>%
  group_by(Region, StationName, Domain)%>%
  select(Region, StationName, Domain, FISH.conc.mn)  %>% 
  spread(Domain, FISH.conc.mn)  #%>%

#Correlation test
set.seed(24)
counts_counts_corr <- as.data.frame(data_4_corr.c) %>% 
  select(-c(Region,StationName))%>% 
  cor_mat(method = "pearson")
#get correlation cooficients
corr_rho.c<- counts_counts_corr  
#check p-values
corr_p.c <- counts_counts_corr %>% cor_get_pval()

#plot
counts_counts_corr %>%
  cor_reorder() %>%
  pull_lower_triangle() %>%
  cor_plot(significant.level = 0.1,label = TRUE)

#write.csv(corr_rho.c, "./corr_table_counts_to_counts.csv")
#write.csv(corr_p.c, "./p_table_counts_to_counts.csv")
```

### Enigmatic microbial lineages increase in cell abundance towards the deep ocean

## Depth profiles by taxa

Select taxa and transform dataframes
```{r vert-corr}
taxa.depth <- c("SAR11", "BACT", "GAM", "ARCH", "CREN", "DELTA", "SAR324",   "SAR202", "SAR406","ALT","POL","ROS","CFX","OPI", "VER","EUB")

#calculate mean conc., adjust the depth
depth_FISH_by_regions <- counts_all%>% ungroup() %>% 
  filter(Domain %in% taxa.depth)%>% 
as.data.frame()

depth_FISH_by_regions_wide<- depth_FISH_by_regions %>%
  group_by(Region, StationName, Depth, Domain) %>% 
  select(Region, StationName, Depth, Domain, FISH.conc.mn)  %>% 
  spread(Domain, FISH.conc.mn)%>% 
  as.data.frame()
```

Correlation between the env parameters
```{r coENV-verr}
set.seed(23)
depth_corr <- depth_FISH_by_regions_wide %>% 
  select(taxa.depth)%>% 
   cor_mat(method = "pearson")
#correlation cooficients
depth_corr_rho <- depth_corr %>% select(rowname,taxa.depth)
#p values
depth_corr_p <- depth_corr %>% cor_get_pval()%>%  select(rowname,taxa.depth)

#plot
if(require("ggpubr")){
  my.palette <- get_palette("PuOr", 200)
  depth_corr %>%
  cor_reorder() %>%
  pull_lower_triangle() %>%
  cor_plot(label = TRUE, palette = my.palette, method = "number")
}
#write.csv(depth_corr_rho, "./depth_corr_rho.csv")
#write.csv(depth_corr_p, "./depth_corr_p.csv")
```


```{r vert-prof}
set.seed(24)
counts_all %>% 
  group_by(Domain) %>% 
  kruskal_test(FISH.conc.mn~Depth)

set.seed(25)
Region_FISH_prop_Wilcox <- counts_all   %>%
  group_by(Domain)%>%
  rstatix::wilcox_test(FISH.conc.mn~Depth) %>%
  adjust_pvalue(method = "BH")%>%
  add_significance()
Region_FISH_prop_Wilcox
```
 
## Save image
```{r save2}
#save.image("Data_analysis.RData")
```
